Follow the Flow of Rain and Paint: \break An Output Sensitive Contour Tree Algorithm

### 2014-09-23 paint-contour-tree.tex
more changes

>>>>

We define a \emph{residue} $R_p$ for each $p \in \cU$. Suppose $p$ has children
$q_1, q_2, \ldots, q_k$ in $\cU$. Then $R_p = |H_p| - \sum_i |H_{q_i}|$. 
By definition, $|H_p| = \sum_{q \in \cU_p} R_p$. 
Note that $R_p$ can be negative. 
Now, define $R^+_p = \max(R_p,0)$, and set $W_p = \sum_{q \in \cU_p} R^+_p$.
Observe that $W_p \geq |H_p|$. We also get an approximate converse.

\begin{claim} \label{clm:loss} $|H_p| \geq W_p - 2\sum_{q \in \cU_p} |q| - 2\sum_{q \in \cU_p} \sum_{h \in \cH} |h \cap R_q|$
\end{claim}

\begin{proof} Consider $q \in \cU_p$, and traverse $\pmax(T)$ up to $p$. Suppose we encounter
the paths $q_0 = q, q_1, q_2, \ldots, q_k = p$. The vertices in the residue $R_q$ that 
are removed from $H_p$ are exactly the vertices that lie in these paths. 
We split this contribution into heavy and light paths, observing that the latter
are all unsupported. Hence, we can (crudely) upper bound the number of vertices
removed by $\sum_{q \in \cU_p} |q| + \sum_{q \in \cU_p} \sum_{h \in \cH} |h \cap q|$.
Each vertex can be removed from at most two distinct $R_q$'s, completing the proof.
\end{proof}

The main challenge of the entire proof is bounding the sum of path lengths, which is done next.
We stress that the all the previous work is mostly setup to get to this claim.

\begin{claim} \label{clm:charge} Consider $p \in \cU$ such that for any $q,q' \in \cU_p$
such that $q$ is a parent of $q'$, $W_q \geq (4/3)W_{q'}$. Then $\sum_{q \in \cU_p} |q| \leq W_p/20$.
\end{claim}

\begin{proof} By the definition of light paths, $|q| < \sqrt{|H_q|}/100 \leq \sqrt{|W_q|}/100$.
We prove that $\sum_{q \in \cU_p} \sqrt{|W_q|}/100 \leq W_p/20$
by a charge redistribution scheme. Assume that each $q \in \cU_p$
starts with $\sqrt{W_p}/100$ units of charge. We redistribute this charge over all
nodes in $\cU_p$, and then calculate the total charge. For $q \in \cU_p$, spread
its charge to all nodes in $\cU_q$ proportional to $R^+$ values. In other words,
give $(\sqrt{W_q}/100)\cdot(R^+_{q'}/W_q)$ units of charge to each $q' \in \cU_q$.

After the redistribution, let us compute the charge deposited at $q$. Every ancestor in $\cU_p$.
$q = a_0, a_1, a_2, \ldots, a_k$ contributes to the charge at $q$. The charge is expressed in the following
equation. We use the assumption that $W_{a_i} \geq (4/3) W_{a_{i-1}}$ and hence $W_{a_i} \geq (4/3)^i$.
$$ ({R^+_q}/100)\sum_{a_i} 1/\sqrt{W_{a_i}} \leq ({R^+_q}/100)\sum_{a_i} (3/4)^{i/2} \leq R^+_q/20 $$
The total charge is $\sum_{q \in \cU_p} R^+_p/20 = W_p/20$.
\end{proof}

We wrap it all up by proving part (ii) of \Lem{cu-root}.

\begin{claim} $\sum_{p \in \cU} |p| \leq H_{r(\cU)}/10$.
\end{claim}

\begin{proof} We use $r$ for $r(\cU)$. Our aim is to simply apply \Clm{charge} from $r(\cU)$. 
Suppose $W_q \geq (4/3)W_{q'}$ (for $q$ parent of $q'$). By \Clm{charge}, $\sum_{p \in \cU} |p| \leq W_{r}/20$.
By \Clm{loss}, $|H_r| \geq W_r - 2\sum_{p \in \cU} |p| - 2\sum_{h \in \cH} |h \cap H_p|$.

We first need to prove the claim assumption,
namely that of $W_q \geq (4/3)W_{q'}$ (for $q$ parent of $q'$). We do know from \Lem{geometric}
that $|H_q| \geq (3/2)|H_{q'}|$. We will apply \Clm{loss} and \Clm{charge} to prove this.

Pick the node furthest from $r(\cU)$
